---
title: "R Basics"
output: html_notebook
---

## The Tidyverse

The tidyverse is a weird name for a collection of user-written packages and functions that really are core to why people use R. These packages are nice in that they share a common format for commands. Click the "play" button in the top right of the gray box below to install the tidyverse on your own computer, if it's not already.

```{r}
# This syntax is common for R --
# Parameters for a command are included in parentheses.
install.packages('tidyverse')

# Note that the hash tag indicates a comment.
```

## Loading Data

There are a few different ways to load data into R. The code below gives some examples. First though, you'll need to set the working directory. If you're doing this in an R project file this is automatic. If not, use the following command:

```{r}
setwd("C:/Users/andre/Box/LearningR") # Change the path to your location
```

Now that R knows where all our files are we can get started. By itself, base R is pretty dumb. It doesn't know how to read many file types (outside of a `.csv`). We need to load some of the packages we installed from the tidyverse before we can access `.dta` files or `.xlsx` files.

### Loading STATA files

There are two ways to load those packages. The first is to type the name of the package before each command. For example, to read a `.dta` file using the haven package, you'd type:

`haven::read_dta('path to the file')`.

It's kinda a pain to have to type the package name before every command though. Instead we'll load the package into the work space.

```{r}
library('haven') # Loads the haven packages to read/write .dta files.

# Here I load a file hosted on github, but you could also just have a file path.

scholarship <- read_dta("https://raw.github.com/andrewmcamp/LearningR/main/scholarship.dta")

```

Let's take a quick look at our data. There are a few options:

```{r}
names(scholarship) # Names just lists the variable names.
```

```{r}
head(scholarship) # Head displays the first 6 rows.
```

That's nice, but how do we get rid of that data set now?

```{r}
ls() # This command lists things in the workspace.

rm(list=ls()) # So let's feed that list into the remove (rm) function.
```

### Loading Package Data

Packages also have data. This can be super handy. Some packages, like `tidycensus` allow you to easily pull census data into your project. Other ones, [like the Urban Institute's education data package](https://github.com/UrbanInstitute/education-data-package-r), let you query and pull in data from IPEDS, NCES Common Core of Data, and others.

I'll be using some California school data from the `AER` package for the rest of the examples.

```{r}
# Install the AER package if you haven't already.
#install.package('AER')

library(AER)
# We can now access the data included with the AER package.
data('CASchools')

```

## Simple Descriptive Statistics

Let's take a closer look at this data

```{r}
summary(CASchools)
```

That's nice but I just want one variable. I can specify the SAT score variable in the scholarship dataframe using the `$` sign and then typing the name of the column.

```{r}
summary(CASchools$expenditure) # dataframe$variable
```

What about several variables?

```{r}
summary(CASchools %>%
  select(english, read, math))
```

The pipe operator `%>%` is a really useful tool. It can be used to string together several functions/arguments and combine several steps into one.

Let's store those test scores separately -- they might be interesting on their own.

```{r}
CATests <- CASchools %>%
  select(english, read, math)

```

Those summary stats are pretty ugly though. Let's use the `stargazer` package to make it nice.

```{r}
stargazer::stargazer(CASchools, type='text', median=TRUE, digits=2, title="CA School Data Set")
```

Notice that this table is missing a few variables? Factor variables (district) and strings are not included in stargazer tables by default (there might be an option).

Stargazer can also output nice HTML and $\LaTeX$ tables. Look up how in the documentation

## Ordinary Least Squares

Alright. Summary Stats. Cool. How about OLS? Basic OLS is referred to as a linear model `lm` in R.

To calculate robust standard errors you need to install the sandwich package if you haven't already. You don't need to load it each time though. R has a few different ways to calculate robust standard errors. I don't know the differences between them, but I know the option `HC1` replicates what STATA does.

Let's regress the school's percent FRL on the school's average math test score.

```{r}
#install.packages('sandwich')

reg1 <- lm(math ~ lunch, data=CASchools)
summary(reg1, robust="HC1") #There are different ways to calculate the robust standard errors. 
#HC1 replicates what Stata does. 
```

What if we wanted to control for student:teacher ratio? That's not included in the dataframe we have...

```{r}
# It's easy to create a new variable.
CASchools$ratio <- CASchools$students / CASchools$teachers


# It's also kind of annoying to keep typing "CASchools$" - Let's do this instead:
attach(CASchools)
```

Now we don't need to specify which dataframe our data is coming from.

```{r}
reg2 <- lm(math ~ lunch + ratio)
summary(reg2, robust="HC1")
```

Neat. Is there some way to visualize this easily. You bet there is. The `ggplot2` and `jtools` package are perfect for making some nice plots. Let's also see an easier way to load packages.

```{r}
pkgs <- c("ggplot2","jtools") # The c operator makes a list of packages
lapply(pkgs,require,character.only=TRUE) # And lapply applies to the list


effect_plot(reg1, pred=lunch, interval = TRUE, plot.points=TRUE)


```

And our second regression

```{r}
effect_plot(reg2, pred=ratio, interval = TRUE, plot.points=TRUE, colors="blue")

```

## Links to useful stuff

[Richard Blissett & Benjamin Skinner - R and STATA side-by-side: Basic commands and processes](http://rslblissett.com/wp-content/uploads/2016/09/sidebyside_130826.pdf)

[Matthieu Gomez - R for STATA users](https://www.matthieugomez.com/statar/)

[RStudio - Cheat sheets](https://rstudio.com/resources/cheatsheets/#contributed-cheatsheets)

[Roger Peng - Intro to the `dplyr` R package](https://www.youtube.com/watch?v=aywFompr1F4)

[William Sundstrom & Michael Kevane - Guide to R: Data analysis for economics (includes code and data)](https://www.dropbox.com/sh/9pz2xtvd31mi4cl/AAAzG8g0D1YVByu7Mrxztxrga?dl=0)
